Thread Pool in GeckoDB/BOLSTER - Integrating a Thread Pool into a high performance Database System \\

### 2018-06-04 Presentations/Paper/paper_thread_pool.tex
Continue writing implementation section of paper

Each assigned task contains a function pointer to a routine and a pointer to data. We implemented the task queue as a generic priority queue that uses a mutex thread to enqueue tasks threadsafe. Every time a new task is enqueued to the task queue, two scenarios can occur. If the new task has a priority, it is placed in the right slot of the task queue regarding its priority and the subsequent tasks are relocated. If the  new task has no priority, it is appended to the end of the queue. Furthermore, tasks can refer to a handle, which is used by the thread pool to wait for a specific amount of tasks.

\subsection{Processing Tasks}
\textbf{TODO: maybe table of provided functions} \\
We define the processing of tasks as a procedure of 3-4 steps. In Figure 2, we show the flow of task enqueueing and execution in the thread pool system. In the first step, tasks are passed as a parameter to the enqueue function provided by the thread pool. The tasks have to be created before passing and later on freed after processing by the calling function.  We provide different functions to enqueue a single task or an array of tasks.

@@

\subsection{Waiting for Tasks}
\begin{figure}
	\includegraphics[width=0.4\textwidth]{img/waitingconcept.png}
	\caption{Concept of waiting for tasks using a slotmap}
	\label{fig2}
\end{figure}
As mentioned in the previous section, ids are assigned to the tasks before they are enqueued.

>>>>

Each assigned task contains a function pointer to a routine and a pointer to data. We implemented the task queue as a generic priority queue that uses a mutex thread to enqueue tasks threadsafe. Every time a new task is enqueued to the task queue, two scenarios can occur. If the new task has a priority, it is placed in the right slot of the task queue regarding its priority and the subsequent tasks are relocated. If the  new task has no priority, it is appended to the end of the queue. Furthermore, tasks can refer to a handle, which is used by the thread pool to wait for a specific amount of tasks.
\begin{table*}[htbp]
	\caption{Functions provided by the thread pool to process tasks}
	\begin{center}
		\begin{tabular}{ c c }
			\hline
			\textbf{Function Name}&\textbf{Explanation}\\
			\hline
			thread\_pool\_enqueue\_task & Enqueue task with optional handle \\
			thread\_pool\_enqueue\_tasks & Enqueue tasks with optional handle \\
			thread\_pool\_wait\_for\_task & Waits until the tasks referenced by the handle are completed. \\

			thread\_pool\_enqueue\_tasks\_wait & Enqueue tasks and wait until they are finished. The main thread also participates in task execution\\ 
			thread\_pool\_wait\_for\_all & Wait until all enqueued tasks are finished. The main thread also participates in task execution \\ \hline
		\end{tabular}
		\label{tab1}
	\end{center}
\end{table*}
\subsection{Processing Tasks}
Processing tasks in a thread pool is a balancing act between generalization the of tasks and the performance of their execution. Therefore, we provide a set of functions to process task in the thread pool to match different requirements of BOLSTER. In Table 1, we show the functions provided by our thread pool to process tasks. One of the requirements of BOLSTER is waiting until a set of tasks is finished. To achieve the same performance as the baseline implementation reaches, we let the main thread first participate in the execution of the tasks and then wait until all tasks with the same handle are finished.

To show the processing of tasks, we consider the function \emph{thread\_pool\_enqueue\_tasks\_wait}. In Figure 2, we show the flow of task enqueueing and execution in the thread pool system using this function. In the first step, tasks are passed as a parameter to the enqueue function provided by the thread pool. The tasks have to be created before passing and later on freed after processing by the calling function.  We provide different functions to enqueue a single task or an array of tasks.

@@

\subsection{Waiting for Tasks}

\begin{figure}
	\includegraphics[width=0.4\textwidth]{img/waitingconcept.png}
	\caption{Concept of waiting for tasks using a slotmap}
	\label{fig2}
\end{figure}
